Proposal outline/notes

### 2012-06-15 docs/Jesse/Proposal/NSNotes.tex
put together rough proposal with a few missing sections

\documentclass{article}

\usepackage{array}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{undertilde}
\usepackage[colorlinks = true, filecolor = red, urlcolor = blue, linkcolor = black]{hyperref}
\usepackage{pdflscape}
\usepackage{pifont}
%\usepackage{fullpage}
\setlength\textwidth{6in}
\setlength\textheight{8in}
\setlength\oddsidemargin{0.25in} % LaTeX adds a default 1in to this!
\setlength\evensidemargin{0.25in}
\setlength\topmargin{-0.0in} % LaTeX adds a default 1in to this!
\setlength\headsep{0in}
\setlength\headheight{0in}
\setlength\footskip{1in}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\vect{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\grad}{\nabla}
\renewcommand{\div}{\grad \cdot}
\newcommand{\ip}[1]{\left\langle #1 \right\rangle}
\newcommand{\eip}[1]{a\left( #1 \right)}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}
\newcommand{\nor}[1]{\left\| #1 \right\|}

\newcommand{\circone}{\ding{192}}
\newcommand{\circtwo}{\ding{193}}
\newcommand{\circthree}{\ding{194}}
\newcommand{\circfour}{\ding{195}}
\newcommand{\circfive}{\ding{196}}

\newcommand{\Reyn}{\rm Re}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Oh}{\Omega_h}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vecttwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vectthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}
\date{}
\author{Jesse Chan}
\title{NS notes}

\begin{document}

@@

\subsection{Mass equation}
\[
\div \vecttwo{\rho u }{\rho v} = 0
\]
\subsection{Momentum equations}
\begin{align*}
\div \vecttwo{\rho u^2+p }{\rho u v} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u v}{\rho v^2+p } &= \div\left(\vec{\sigma_{i2}}\right)
\end{align*}
\subsection{Energy equation}
Let $\boldsymbol \sigma$ be a tensor whose $ij$th term is $\sigma_{ij}$.  
\[
\div\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} = \div\left[\boldsymbol\sigma U + \vec{q}\right]
\]
\subsection{Newtonian fluid laws}
We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
%\[
%\epsilon_{ij} = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
%\] 
%which translates to
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
\]
Thus our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the symmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

\subsection{Fourier's heat conduction law}
We assume Fourier's law 
\begin{align*}
\vec{q} &= \kappa \grad T,
\end{align*}
where $\kappa$ is generally a function of temperature. We introduce here the Prandtl number here as well
\[
{\rm Pr} = \frac{\gamma c_v \mu}{\kappa}.
\]
In this case, we assume a constant Prandlt number, which implies that the heat conductivity $\kappa$ is proportional to viscosity $\mu$.

\subsection{Classical and momentum variables}

In classical variables, our system of equations is now
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol\sigma \mathbf{u} + \vec{q}\right) &=0\\
\frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I &= \grad \mathbf{u} - \boldsymbol \omega\\
\frac{1}{\kappa}\vec{q} &= \grad T
\end{align*}

@@

The test norm adopted for convection-diffusion in \cite{DPGrobustness,DPGrobustness2} is defined elementwise on $K$ as
\[
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2.
\]
This test norm both delivers robust control of the error in the $L^2$ variables $u$ and $\sigma$ and avoids boundary layers in the computation of local test functions.

@@

\bibliographystyle{plain}
\bibliography{notes}

\end{document}

>>>>

@@

\begin{itemize}
\item \textbf{Mass equation}
\[
\div \vecttwo{\rho u }{\rho v} = 0
\]
\subsection{Momentum equations}
\begin{align*}
\div \vecttwo{\rho u^2+p }{\rho u v} &= \div\left(\vec{\sigma_{i1}}\right)\\
\div \vecttwo{\rho u v}{\rho v^2+p } &= \div\left(\vec{\sigma_{i2}}\right)
\end{align*}

\item \textbf{Energy equation}
Let $\boldsymbol \sigma$ be a tensor whose $ij$th term is $\sigma_{ij}$.  
\[
\div\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} = \div\left[\boldsymbol\sigma U + \vec{q}\right]
\]

\item \textbf{Newtonian fluid laws}
We represent $\boldsymbol\sigma$ using the Newtonian fluid law
\[
\sigma_{ij} = \mu(u_{i,j} + u_{j,i}) + \lambda u_{k,k} \delta_{ij}
\]
where $\mu$ is viscosity and $\lambda$ is bulk viscosity. 
%Following elasticity, a general formula for the stress tensor is 
%\[
%\sigma_{ij} = 2\mu \epsilon_{ij} + \lambda \epsilon_{kk} \delta_{ij}
%\]
We can invert the stress tensor under isotropic and plane strain assumptions to get
%\[
%\epsilon_{ij} = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
%\] 
%which translates to
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \frac{1}{2\mu} \sigma_{ij} - \frac{\lambda}{4\mu (\mu + \lambda)} \sigma_{kk}\delta_{ij}
\]
We also have
\[
\frac{1}{2}\left(\grad  U + \grad ^T  U\right) = \grad  U - \boldsymbol \omega
\]
where $\boldsymbol \omega$ is the antisymmetric part of the infinitesimal strain tensor:
\[
\boldsymbol \omega = \frac{1}{2}\left(\grad  U - \grad ^T  U\right).
\]
Thus our final form is
\begin{align*}
\grad  U - \boldsymbol \omega = \frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I.
\end{align*}
Notice that $\boldsymbol \omega$ is implicitly defined to be the symmetric part of $\grad u$ by taking the symmetric part of the above equation.

@@

\item \textbf{Fourier's heat conduction law}
We assume Fourier's law 
\begin{align*}
\vec{q} &= \kappa \grad T,
\end{align*}
where $\kappa$ is generally a function of temperature. We introduce here the Prandtl number here as well
\[
{\rm Pr} = \frac{\gamma c_v \mu}{\kappa}.
\]
In this case, we assume a constant Prandlt number, which implies that the heat conductivity $\kappa$ is proportional to viscosity $\mu$.

To summarize, our system of equations in the classical variables is now
\begin{align*}
\div \vecttwo{\rho u }{\rho v} &= 0\\
\div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0\\
\div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol\sigma \mathbf{u} + \vec{q}\right) &=0\\
\frac{1}{2\mu} \boldsymbol \sigma - \frac{\lambda}{4\mu (\mu + \lambda)} { \rm tr}(\boldsymbol \sigma) \boldsymbol I &= \grad \mathbf{u} - \boldsymbol \omega\\
\frac{1}{\kappa}\vec{q} &= \grad T
\end{align*}

@@

The test norm adopted for convection-diffusion in Section~\ref{blah}\todo{get section} and in \cite{DPGrobustness2} is defined elementwise on $K$ as
\[
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2.
\]
This test norm both delivers robust control of the error in the $L^2$ variables $u$ and $\sigma$ and avoids boundary layers in the computation of local test functions.

@@
